---
title: "Groups on Russia Preliminary Report"
author: "Rafael Campos-Gottardo & Alan Nemirovski"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("scripts/2_group_variables.R")
```

# Method 1: Group-Based Indices

## Creating Group-Based Indices

**Pro Russian Long Term**

0 - 6 scale

Respondents gain 1 point if: - They do not think Russia is their countries greatest threat (Q68). - They think that Europeans should pursue trade and diplomacy with Russia (Q73). - They think their country spends too much on defence (Q69) - They do not support democracy (Q5 \< 5/10) - They do not trust the European Union (New_Q43)

```{r}
hist(EUI_data$Pro_russia_scale)
```

**Pro Russian Short Term**

0 - 7 scale

-   Respondents gain 1 point if:
    -   They do not think Russia is their countries greatest threat (Q68).
    -   They think that Europeans should pursue trade and diplomacy with Russia (Q73).
    -   They think their country spends too much on defence (Q69)
    -   They do not support democracy (Q5 \< 5/10)
    -   They do not trust the European Union (New_Q43)
    -   They do want Russia to gain territory in Ukraine (EUI_Ukraine_Outcome)
    -   They do not support Aid to Ukraine (New_Q78, less than 4 on a 0-9 scale)

Question: Q5 only captures ideological pro Russia not situational.

```{r}
hist(EUI_data$Pro_russia_scale_short)
```

**Russian Appeaser Long Term**

0 - 6 scale

-   Respondents gain 1 point if:
    -   they think Russia is their country's greatest threat (Q68) or they think Nuclear proliferation is their greatest threat.
    -   They want to trade with Russia (instead of invest in defence) (Q73)
    -   They think their country spends too much on defence (Q69)
    -   They oppose the creation of an integrated European army (Q70)
    -   They think that NATO is not important for their country's defence (Q71)
    -   They would vote to remain in the EU (Q9)

```{r}
hist(EUI_data$appease_russia_scale)
```

**Russian Appeaser Short Term**

Only post 2023

0 - 8 scale

Respondents gain 1 point if: - They do not think Russia is their country's greatest threat (Q68) - They want to trade with Russia (instead of invest in defence) (Q73) - Their country spends too much on defence (Q69) - They oppose the creation of an integrated European army (Q70) - They think that NATO is not important for their country's defence (Q71) - They would vote to remain in the EU (Q9) - Favour a return to the territorial situation before the war (EUI_Ukraine_Outcome) - Support only compromises to end the war (New_Q78_variables)

```{r cars}
hist(EUI_data$appease_russia_scale_short)
```

## Testing Method 1 Variables

### Correlations

First, I check the correlation of the pro-Russia and appease Russia scales to ensure they are capturing different concepts.

The long term scales

```{r}
cor(EUI_data$Pro_russia_scale, EUI_data$appease_russia_scale, use = "complete.obs")
```

The short term scales

```{r}
cor(EUI_data$Pro_russia_scale_short, EUI_data$appease_russia_scale_short,
    use = "complete.obs")
```

Both scales are weakly correlated at $r = 0.27$ and $r = 0.35$ respectively. Indicating that the scales are not highly collinear.

# Method 2: Combined index

## Generating combined index

*Long term appeaser-supporter scale*

0 - 7 scale

Respondents get 1 point if:

-   They do not think Russia or nuclear non-proliferation is their countries greatest threat (Q68).

-   They think that Europeans should pursue trade and diplomacy with Russia (Q73).

-   They think their country spends too much on defence (Q69)

-   They do not support democracy (Q5 \< 5/10)

-   They do not trust the European Union (New_Q43)

Respondents get 0.5 points if:

-   **\*INSERT HERE\***

```{r}
hist(EUI_data$comb_scale)
```

*Short term appeaser-supporter scale* 0 - 9.5 scale

Respondents get 1 point if:

-   They do not think Russia is their countries greatest threat (Q68).

-   They think that Europeans should pursue trade and diplomacy with Russia (Q73).

-   They think their country spends too much on defence (Q69)

-   They do not support democracy (Q5 \< 5/10)

-   They do not trust the European Union (New_Q43)

Respondents get an additional 0.5 point if:

-   **\*\*INSERT\*\***

```{r}
hist(EUI_data$comb_scale_short)
```

# Comparisons of measures

### Over time distributions

Long term pro-Russia

```{r}
EUI_data %>% 
  filter(Year > 2019) %>% 
  ggplot(aes(x = Pro_russia_scale, fill = as.factor(Year), group = as.factor(Year))) + 
  geom_bar(position = "dodge") + 
  labs(fill = "Year", x = "Long Term Pro-Russia Scale") + 
  scale_x_continuous(breaks = 0:6) +
  theme_bw()

```

Short term pro-Russia

```{r}
EUI_data %>% 
  filter(Year > 2022) %>% 
  ggplot(aes(x = Pro_russia_scale, fill = as.factor(Year), group = as.factor(Year))) + 
  geom_bar(position = "dodge") + 
  labs(fill = "Year", x = "Short Term Pro-Russia Scale") + 
  scale_x_continuous(breaks = 0:7) +
  theme_bw()
```

Long term Appease Russia

```{r}
EUI_data %>% 
  filter(Year > 2019) %>% 
  ggplot(aes(x = appease_russia_scale, fill = as.factor(Year), group = as.factor(Year))) + 
  geom_bar(position = "dodge") + 
  scale_x_continuous(breaks = 0:6) +
  labs(fill = "Year", x = "Long Term Appease Russia Scale") + 
  theme_bw()
```

Short term Appease Russia

```{r}

EUI_data %>% 
  filter(Year > 2019) %>% 
  ggplot(aes(x = appease_russia_scale_short, fill = as.factor(Year), group = as.factor(Year))) + 
  geom_bar(position = "dodge") + 
  scale_x_continuous(breaks = 0:8) +
  labs(fill = "Year", x = "Short Term Appease Russia Scale") + 
  theme_bw()

```

Long term Combined Scale

```{r}
EUI_data %>% 
  filter(Year > 2019) %>% 
  ggplot(aes(x = comb_scale, fill = as.factor(Year), group = as.factor(Year))) + 
  geom_bar(position = "dodge") + 
  scale_x_continuous(breaks = 0:8) +
  labs(fill = "Year", x = "Long Term Combined Scale") + 
  theme_bw()
```

Short term Combined Scale

```{r}
EUI_data %>% 
  filter(Year > 2019) %>% 
  ggplot(aes(x = comb_scale_short, fill = as.factor(Year), group = as.factor(Year))) + 
  geom_bar(position = "dodge") + 
  scale_x_continuous(breaks = 0:8) +
  labs(fill = "Year", x = "Short Term Combined Scale") + 
  theme_bw()
```

### Average Measures (ALL) by Country

Long Term

```{r}
EUI_data %>% 
  filter(Year > 2019) %>% 
  group_by(country) %>% 
  summarise(`Pro-Russian Scale` = mean(Pro_russia_scale, na.rm = TRUE),
            `Appease Russia Scale` = mean(appease_russia_scale, na.rm = TRUE),
            `Combined Scale` = mean(comb_scale, na.rm = TRUE)) %>% 
    arrange(-`Pro-Russian Scale`) %>% 
  kable(digits = 2) 

```

Short Term

```{r}
EUI_data %>% 
  filter(Year > 2019) %>% 
  group_by(country) %>% 
  summarise(`Pro-Russian Scale` = mean(Pro_russia_scale_short, na.rm = TRUE),
            `Appease Russia Scale` = mean(appease_russia_scale_short, na.rm = TRUE),
            `Combined Scale` = mean(comb_scale_short, na.rm = TRUE)) %>% 
    arrange(-`Pro-Russian Scale`) %>% 
  kable(digits = 2) 
```

## Create group variables

(Using Pro-Russia + Russia Appeaser scales:) Those who score 3 or less on both scales are considered pro-Europeans, those who score 4 or above on the Russian appeaser scale are considered Russian appeasers and those that score 4 or above on the Pro Russia scale are considered Pro-Russian.

```{r}
EUI_data <- EUI_data %>% 
  mutate(Russia_groups_long = case_when(appease_russia_scale < 4 & Pro_russia_scale < 4 ~ "Anti-Russia",
                                        appease_russia_scale > 3 ~ "Russia Appeaser",
                                        Pro_russia_scale > 3 ~ "Pro Russian"),
         Russia_groups_short = case_when(appease_russia_scale_short < 4 & Pro_russia_scale_short < 4 ~ "Anti-Russia",
                                        appease_russia_scale_short > 3 ~ "Russia Appeaser",
                                        Pro_russia_scale_short > 3 ~ "Pro Russian")
         )

```

*Note: A similar grouping measure could be reconstructed with thresholds set on the combined index.*

### Groups over time

Long Term Variable

```{r}
EUI_data %>% 
  filter(Year > 2019) %>% 
  group_by(Year, Russia_groups_long) %>% 
  summarise(n = n(), .groups = "drop_last") %>% 
  filter(!is.na(Russia_groups_long)) %>% 
  mutate(Percentage = n / sum(n) * 100) %>% 
  kable(digits = 2)
```

Short Term Variable

```{r}
EUI_data %>% 
  filter(Year > 2022) %>% 
  group_by(Year, Russia_groups_short) %>% 
  summarise(n = n(), .groups = "drop_last") %>% 
  filter(!is.na(Russia_groups_short)) %>% 
  mutate(Percentage = n / sum(n) * 100) %>% 
  kable(digits = 2)
```

### Groups by Country

Long Term Measure

```{r}
EUI_data %>% 
  filter(Year > 2019) %>% 
  group_by(country, Russia_groups_long) %>% 
  summarise(n = n(), .groups = "drop_last") %>% 
  filter(!is.na(Russia_groups_long)) %>% 
  mutate(Percentage = n / sum(n) * 100) %>% 
  kable(digits = 2)
```

Short Term Measure

```{r}
EUI_data %>% 
  filter(Year > 2022) %>% 
  group_by(country, Russia_groups_short) %>% 
  summarise(n = n(), .groups = "drop_last") %>% 
  filter(!is.na(Russia_groups_short)) %>% 
  mutate(Percentage = n / sum(n) * 100) %>% 
  kable(digits = 2)
```
