---
title: "Groups on Russia Preliminary Report"
author: "Rafael Campos-Gottardo & Alan Nemirovski"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("scripts/2_group_variables.R")
```

# Updated Scale based on Dietlind's Comments Feb 4, 2026

Slimmed down indicators 
Long Term
Group 1: Not threat, trade and diplomacy
Group 2: Threat, trade and diplomacy
Group 3: Threat, Troops?, military 
Other: don't know 

```{r}

 EUI_data <- EUI_data %>% 
  mutate(Ukraine_group_long = case_when((Q68 == 2 | Q68 == 5) & Q73 == "European countries should invest more in trade and diplomacy with Russia to improve relations" ~ "Russia Appeaser",
    Q68 != 2 & Q73 == "European countries should invest more in trade and diplomacy with Russia to improve relations" ~ "Pro-Russia",
                                       Q73 == "European countries should invest more in defence and security to defend against Russian aggression" ~ "Pro-Ukraine",
                                       TRUE ~ "Indifferent"
                                   ))


EUI_2023 <- EUI_2023 %>% 
  mutate(Ukraine_group_2023 = case_when(New_Q79 %in% c(1, 2, 3) &
                                          EUI_Ukraine_Outcome %in% c(1, 2) ~ "Pro-Russia",
                                        
                                         New_Q79 %in% c(4, 5) & EUI_Ukraine_Outcome == 3 &
                                          (New_Q78_1 %in% c(3, 4) | New_Q78_3 %in% c(3,4)) &
                                           (New_Q78_4 %in% c(1, 2) & New_Q78_4 %in% c(1, 2) |
                                              New_Q78_5 %in% c(1, 2)) ~ "Russia Appeaser",
                                        
                                         New_Q79 %in% c(4, 5) & EUI_Ukraine_Outcome %in% c(5, 4) & 
                                           New_Q78_1 %in% c(3, 4) & New_Q78_2 %in% c(3, 4) &
                                           New_Q78_3 %in% c(3, 4) & New_Q78_4 %in% c(3, 4) &
                                           New_Q78_5 %in% c(3, 4) & New_Q78_7 %in% c(3, 4) ~ "Pro-Ukraine",
                                        
                                         TRUE ~ "Indifferenent"
                                           ))

EUI_data <- EUI_data %>% 
  mutate(Ukraine_group_short = case_when( EUI_Ukraine_Outcome %in% c(1, 2) & 
                                           New_Q78_11 %in% c(3, 4) ~ "Pro-Russia",
                                          
                                         EUI_Ukraine_Outcome == 3 &
                                           (New_Q78_1 %in% c(3, 4) | New_Q78_3 %in% c(3,4)) &
                                           (New_Q78_4 %in% c(1, 2) & New_Q78_4 %in% c(1, 2) |
                                              New_Q78_5 %in% c(1, 2)) &
                                           New_Q78_11 %in% c(3, 4) ~ "Russia Appeaser",
                                         
                                          EUI_Ukraine_Outcome %in% c(5, 4) & 
                                           (New_Q78_1 %in% c(3, 4) | New_Q78_2 %in% c(3, 4) |
                                           New_Q78_3 %in% c(3, 4) | New_Q78_4 %in% c(3, 4) |
                                           New_Q78_5 %in% c(3, 4) |
                                             New_Q78_7 %in% c(3, 4)) ~ "Pro Ukraine",
                                         
                                         TRUE ~ "Indifferenent"
                                           ))

table(EUI_2023$Ukraine_group_2023)
table(EUI_2023$New_Q79)
```

### Long-term group trends

```{r}
EUI_data %>% 
  filter(Year > 2019) %>% 
  ggplot(aes(x = Ukraine_group_long, fill = as.factor(Year), group = as.factor(Year))) + 
  geom_bar(position = "dodge") + 
  labs(fill = "Year", title = "Russia Attitudinal Groups: Long-Term Measure",
       x = "Groups", y = "Frequency") + 
  theme_bw()

```

### Short-term group trends

```{r}
EUI_data %>% 
  filter(Year > 2022) %>% 
  ggplot(aes(x = Ukraine_group_short, fill = as.factor(Year), group = as.factor(Year))) + 
  geom_bar(position = "dodge") + 
  labs(fill = "Year", title = "Russia Attitudinal Groups: Short-Term Measure",
       x = "Groups", y = "Frequency") + 
  theme_bw()
```

## Short Term Measure

New_Q79: (2023: 5923; 14502; 3836) anti Russia 
<1>     	Entirely NATO (Pro Russian)
<2>     	More NATO than Russia (Pro Russian)
<3>     	NATO and Russia equally (Pro Russian)
<4>     	More Russia than NATO (Anti Russian)
<5>     	Entirely Russia (Anti Russian)
<6 fixed>     	Don’t know (Other)

Outcome of the War
Group 1: (7764); Group 2: 17469; Group 3: 32023; 16229
[EUI_Ukraine_Outcome] Generally speaking, what do you consider an ideal end to the war in Ukraine?
<1>     	Russia achieves all its territorial goals in Ukraine
<2>     	Russia manages to take more Ukrainian territory than it controlled at the start of the war
<3>     	A return to the territorial situation of earlier this year before the war started, including Russia holding territory it took from Ukraine in 2014
<4>     	Ukraine manages to retake some of the territory Russia occupied in 2014
<5>     	Ukraine manages to retake all of its original territory (including Crimea)
<99 fixed>     	Don’t know

# Proposal to code 

[New_Q78] Thinking about the conflict in Ukraine, to what extent would you support or oppose each of the following? (Please select the option that best applies on each row)
-[New_Q78_1]     	$Qcountry2 allowing Ukrainian refugees to enter the country (Appease or Pro_Ukraine)
-[New_Q78_3]     	$Qcountry2 sending humanitarian aid to Ukraine (Appease or Pro_Ukraine)
-[New_Q78_4]     	$Qcountry2 sending weapons and military equipment to support the Ukrainian army (Pro Ukraine)
-[New_Q78_5]     	$Qcountry2 accepting higher energy costs due to the sanctions imposed on Russia (Pro Ukraine)
-[New_Q78_7]     	Ukraine fast-tracking European Union membership (Pro Ukraine)
-[New_Q78_11]     	$Qcountry2 facilitating diplomatic relations between Russia and Ukraine (Appease or Pro Russia)

Ukraine Aid Questions


## Long Term Variable 

### Breakdown by country and year pro-Russia

```{r}
europe <- ne_countries(scale = "medium", returnclass = "sf") %>%
  filter(region_un == "Europe")

europe <- europe %>% 
 # st_transform(3035) %>% 
  mutate(name = replace(name, name == "United Kingdom", "UK"))

years <- seq(2020, 2025, 1)
maps_pro_Russia <- list()
for(i in 1:length(years)){
  df <- EUI_data %>% 
    filter(Year == years[i]) %>% 
    mutate(Pro_Russia = case_when(Ukraine_group_long == "Pro-Russian" ~ 1,
                                  TRUE ~ 0)) %>% 
    group_by(country) %>% 
    summarise(`Pro-Russia` = mean(Pro_Russia, na.rm = TRUE))
  
  
  df <- left_join(europe, df, by = c("name" = "country"))
  
  maps_pro_Russia[[as.character(years[i])]] <- df %>% 
    ggplot() + 
    geom_sf(aes(fill = `Pro-Russia`)) + 
    scale_fill_gradientn(
    colours = c("pink","darkred"), # light pink → hot pink → purple
    na.value = "grey90",
    breaks = seq(0, 0.7, 0.1),
    limits = c(0, 0.7),
    labels = scales::percent
  ) +
      coord_sf(
    xlim = c(-25, 49),   # longitude range
    ylim = c(34, 72)     # latitude range
  ) + 
    theme_void() +
      labs(fill = "% Pro-Russia",
           title = paste0(years[i])) + 
  theme(legend.position = "bottom",
        text = element_text(size = 8))
}

ggpubr::ggarrange(plotlist = maps_pro_Russia[c(1, 3, 5, 6)],
  ncol = 2,
  nrow = 2,
  common.legend = TRUE,
  legend = "bottom")
```

## Russia Appeaser

```{r}
maps_appease <- list()
for(i in 1:length(years)){
  df <- EUI_data %>% 
    filter(Year == years[i]) %>% 
    mutate(Appeaser = case_when(Ukraine_group_long == "Appease" ~ 1,
                                  TRUE ~ 0)) %>% 
    group_by(country) %>% 
    summarise(`Russia Appeaser` = mean(Appeaser, na.rm = TRUE))
  
  
  df <- left_join(europe, df, by = c("name" = "country"))
  
  maps_appease[[as.character(years[i])]] <- df %>% 
    ggplot() + 
    geom_sf(aes(fill = `Russia Appeaser`)) + 
    scale_fill_gradientn(
    colours = c("white","blue"), # light pink → hot pink → purple
    na.value = "grey90",
    breaks = seq(0, 0.3, 0.1),
    limits = c(0, 0.3),
    labels = scales::percent
  ) +
      coord_sf(
    xlim = c(-25, 49),   # longitude range
    ylim = c(34, 72)     # latitude range
  ) + 
    theme_void() +
      labs(fill = "% Appease Russia",
           title = paste0(years[i])) + 
  theme(legend.position = "bottom",
        text = element_text(size = 8))
}

ggpubr::ggarrange(plotlist = maps_appease[c(1, 3, 5, 6)],
  ncol = 2,
  nrow = 2,
  common.legend = TRUE,
  legend = "bottom")
```

## Pro Ukraine

```{r}
maps_pro_ukraine <- list()
for(i in 1:length(years)){
  df <- EUI_data %>% 
    filter(Year == years[i]) %>% 
    mutate(Ukraine = case_when(Ukraine_group_long == "EU Defence" ~ 1,
                                  TRUE ~ 0)) %>% 
    group_by(country) %>% 
    summarise(`Pro-Ukraine` = mean(Ukraine, na.rm = TRUE))
  
  
  df <- left_join(europe, df, by = c("name" = "country"))
  
  maps_pro_ukraine[[as.character(years[i])]] <- df %>% 
    ggplot() + 
    geom_sf(aes(fill = `Pro-Ukraine`)) + 
    scale_fill_gradientn(
    colours = c("yellow","orange"), # light pink → hot pink → purple
    na.value = "grey90",
    breaks = seq(0, 0.7, 0.1),
    limits = c(0, 0.7),
    labels = scales::percent
  ) +
      coord_sf(
    xlim = c(-25, 49),   # longitude range
    ylim = c(34, 72)     # latitude range
  ) + 
    theme_void() +
      labs(fill = "% Pro-Ukraine",
           title = paste0(years[i])) + 
  theme(legend.position = "bottom",
        text = element_text(size = 8))
}

ggpubr::ggarrange(plotlist = maps_pro_ukraine[c(1, 3, 5, 6)],
  ncol = 2,
  nrow = 2,
  common.legend = TRUE,
  legend = "bottom")
```

## Ambivalent

```{r}
maps_other <- list()
for(i in 1:length(years)){
  df <- EUI_data %>% 
    filter(Year == years[i]) %>% 
    mutate(Ambivelent = case_when(Ukraine_group_long == "Other" ~ 1,
                                  TRUE ~ 0)) %>% 
    group_by(country) %>% 
    summarise(Ambivelent = mean(Ambivelent, na.rm = TRUE))
  
  
  df <- left_join(europe, df, by = c("name" = "country"))
  
  maps_other[[as.character(years[i])]] <- df %>% 
    ggplot() + 
    geom_sf(aes(fill = Ambivelent)) + 
    scale_fill_gradientn(
    colours = c("lightgreen","darkgreen"), # light pink → hot pink → purple
    na.value = "grey90",
    breaks = seq(0, 0.4, 0.1),
    limits = c(0, 0.4),
    labels = scales::percent
  ) +
      coord_sf(
    xlim = c(-25, 49),   # longitude range
    ylim = c(34, 72)     # latitude range
  ) + 
    theme_void() +
      labs(fill = "% Ambivelent",
           title = paste0(years[i])) + 
  theme(legend.position = "bottom",
        text = element_text(size = 8))
}

ggpubr::ggarrange(plotlist = maps_other[c(1, 3, 5, 6)],
  ncol = 2,
  nrow = 2,
  common.legend = TRUE,
  legend = "bottom")
```


# Method 1: Group-Based Indices

## Creating Group-Based Indices

**Pro Russian Long Term**

0 - 6 scale

Respondents gain 1 point if:
- They do not think Russia is their countries greatest threat (Q68). 
- They think that Europeans should pursue trade and diplomacy with Russia (Q73).
- They think their country spends too much on defence (Q69) - They do not support democracy (Q5 \< 5/10) - They do not trust the European Union (New_Q43)

```{r}
hist(EUI_data$Pro_russia_scale)
```

**Pro Russian Short Term**

0 - 7 scale

-   Respondents gain 1 point if:
    -   They do not think Russia is their countries greatest threat (Q68).
    -   They think that Europeans should pursue trade and diplomacy with Russia (Q73).
    -   They think their country spends too much on defence (Q69)
    -   They do not support democracy (Q5 < 5/10)
    -   They do not trust the European Union (New_Q43)
    -   They do want Russia to gain territory in Ukraine (EUI_Ukraine_Outcome)
    -   They do not support Aid to Ukraine (New_Q78, less than 4 on a 0-9 scale)

Question: Q5 only captures ideological pro Russia not situational.

```{r}
hist(EUI_data$Pro_russia_scale_short)
```

## Biggest Groups 

### Do not think Russia is the biggest threat

```{r}
table(EUI_data$Q68 != 2)
```

### Do not support high levels of aid

```{r}
table(EUI_data$Aid_support < 4)
```

**Russian Appeaser Long Term**

0 - 6 scale

-   Respondents gain 1 point if:
    -   they think Russia is their country's greatest threat (Q68) or they think Nuclear proliferation is their greatest threat.
    -   They want to trade with Russia (instead of invest in defence) (Q73)
    -   They think their country spends too much on defence (Q69)
    -   They oppose the creation of an integrated European army (Q70)
    -   They think that NATO is not important for their country's defence (Q71)
    -   They would vote to remain in the EU (Q9)

```{r}
hist(EUI_data$appease_russia_scale)
```

## Biggest groups 

### they think Russia is their country's greatest threat (Q68) or they think Nuclear proliferation is their greatest threat.
```{r}
table(EUI_data$Q68 %in% c(2, 5))
```

### They would vote to remain in the EU

```{r}
table(EUI_data$Q9i == 1)
```

**Russian Appeaser Short Term**

Only post 2023

0 - 8 scale

Respondents gain 1 point if:
- They do they think Russia is their country's greatest threat (Q68) 
- They want to trade with Russia (instead of invest in defence) (Q73)
- Their country spends too much on defence (Q69)
- They oppose the creation of an integrated European army (Q70)
- They think that NATO is not important for their country's defence (Q71) 
- They would vote to remain in the EU (Q9)
- Favour a return to the territorial situation before the war (EUI_Ukraine_Outcome)
- Support only compromises to end the war (New_Q78_variables)

```{r}
hist(EUI_data$appease_russia_scale_short)
```

## Testing Method 1 Variables

### Correlations

First, I check the correlation of the pro-Russia and appease Russia scales to ensure they are capturing different concepts.

The long term scales

```{r}
cor(EUI_data$Pro_russia_scale, EUI_data$appease_russia_scale, use = "complete.obs")
```

The short term scales

```{r}
cor(EUI_data$Pro_russia_scale_short, EUI_data$appease_russia_scale_short,
    use = "complete.obs")
```

Both scales are weakly correlated at $r = 0.27$ and $r = 0.35$ respectively. Indicating that the scales are not highly collinear.

# Method 2: Combined index

## Generating combined index

*Long term appeaser-supporter scale*

0 - 7 scale

Respondents get 1 point if:

-   They do not think Russia or nuclear non-proliferation is their countries greatest threat (Q68).

-   They think that Europeans should pursue trade and diplomacy with Russia (Q73).

-   They think their country spends too much on defence (Q69)

-   They do not support democracy (Q5 \< 5/10)

-   They do not trust the European Union (New_Q43)

Respondents get 0.5 points if:

- Do not support an integrated European army (Q70). 
- They do not think NATO is important to their country's defence (Q71).
- They would vote to remain in the EU (Q9)

```{r}
hist(EUI_data$comb_scale)
```

*Short term appeaser-supporter scale* 0 - 9.5 scale

Respondents get 1 point if:

-   They do not think Russia is their countries greatest threat (Q68).

-   They think that Europeans should pursue trade and diplomacy with Russia (Q73).

-   They think their country spends too much on defence (Q69)

-   They do not support democracy (Q5 \< 5/10)

-   They do not trust the European Union (New_Q43)

Respondents get 0.5 points if:

- Do not support an integrated European army (Q70). 
- They do not think NATO is important to their country's defence (Q71).
- They would vote to remain in the EU (Q9)
- They support a return to the pre 2022 status quo in Ukraine (EUI_Ukraine_Outcome).
- Support only compromises to end the war (New_Q78_variables)

```{r}
hist(EUI_data$comb_scale_short)
```

# Comparisons of measures

### Over time distributions

Long term pro-Russia

```{r}
EUI_data %>% 
  filter(Year > 2019) %>% 
  ggplot(aes(x = Pro_russia_scale, fill = as.factor(Year), group = as.factor(Year))) + 
  geom_bar(position = "dodge") + 
  labs(fill = "Year", x = "Long Term Pro-Russia Scale") + 
  scale_x_continuous(breaks = 0:6) +
  theme_bw()

```

Short term pro-Russia

```{r}
EUI_data %>% 
  filter(Year > 2022) %>% 
  ggplot(aes(x = Pro_russia_scale, fill = as.factor(Year), group = as.factor(Year))) + 
  geom_bar(position = "dodge") + 
  labs(fill = "Year", x = "Short Term Pro-Russia Scale") + 
  scale_x_continuous(breaks = 0:7) +
  theme_bw()
```

Long term Appease Russia

```{r}
EUI_data %>% 
  filter(Year > 2019) %>% 
  ggplot(aes(x = appease_russia_scale, fill = as.factor(Year), group = as.factor(Year))) + 
  geom_bar(position = "dodge") + 
  scale_x_continuous(breaks = 0:6) +
  labs(fill = "Year", x = "Long Term Appease Russia Scale") + 
  theme_bw()
```

Short term Appease Russia

```{r}

EUI_data %>% 
  filter(Year > 2019) %>% 
  ggplot(aes(x = appease_russia_scale_short, fill = as.factor(Year), group = as.factor(Year))) + 
  geom_bar(position = "dodge") + 
  scale_x_continuous(breaks = 0:8) +
  labs(fill = "Year", x = "Short Term Appease Russia Scale") + 
  theme_bw()

```

Long term Combined Scale

```{r}
EUI_data %>% 
  filter(Year > 2019) %>% 
  ggplot(aes(x = comb_scale, fill = as.factor(Year), group = as.factor(Year))) + 
  geom_bar(position = "dodge") + 
  scale_x_continuous(breaks = 0:8) +
  labs(fill = "Year", x = "Long Term Combined Scale") + 
  theme_bw()
```

Short term Combined Scale

```{r}
EUI_data %>% 
  filter(Year > 2019) %>% 
  ggplot(aes(x = comb_scale_short, fill = as.factor(Year), group = as.factor(Year))) + 
  geom_bar(position = "dodge") + 
  scale_x_continuous(breaks = 0:8) +
  labs(fill = "Year", x = "Short Term Combined Scale") + 
  theme_bw()
```

### Average Measures (ALL) by Country

Long Term

```{r}
EUI_data %>% 
  filter(Year > 2019) %>% 
  group_by(country) %>% 
  summarise(`Pro-Russian Scale` = mean(Pro_russia_scale, na.rm = TRUE),
            `Appease Russia Scale` = mean(appease_russia_scale, na.rm = TRUE),
            `Combined Scale` = mean(comb_scale, na.rm = TRUE)) %>% 
    arrange(-`Pro-Russian Scale`) %>% 
  kable(digits = 2) 

```

Short Term

```{r}
EUI_data %>% 
  filter(Year > 2019) %>% 
  group_by(country) %>% 
  summarise(`Pro-Russian Scale` = mean(Pro_russia_scale_short, na.rm = TRUE),
            `Appease Russia Scale` = mean(appease_russia_scale_short, na.rm = TRUE),
            `Combined Scale` = mean(comb_scale_short, na.rm = TRUE)) %>% 
    arrange(-`Pro-Russian Scale`) %>% 
  kable(digits = 2) 
```

## Create group variables

(Using Pro-Russia + Russia Appeaser scales:) Those who score 3 or less on both scales are considered pro-Europeans, those who score 4 or above on the Russian appeaser scale are considered Russian appeasers and those that score 4 or above on the Pro Russia scale are considered Pro-Russian.

```{r}
EUI_data <- EUI_data %>% 
  mutate(Russia_groups_long = case_when(appease_russia_scale < 4 & Pro_russia_scale < 4 ~ "Anti-Russia",
                                        appease_russia_scale > 3 ~ "Russia Appeaser",
                                        Pro_russia_scale > 3 ~ "Pro Russian"),
         Russia_groups_short = case_when(appease_russia_scale_short < 4 & Pro_russia_scale_short < 4 ~ "Anti-Russia",
                                        appease_russia_scale_short > 3 ~ "Russia Appeaser",
                                        Pro_russia_scale_short > 3 ~ "Pro Russian")
         )

```

*Note: A similar grouping measure could be reconstructed with thresholds set on the combined index.*

### Groups over time

Long Term Variable

```{r}
EUI_data %>% 
  filter(Year > 2019) %>% 
  group_by(Year, Russia_groups_long) %>% 
  summarise(n = n(), .groups = "drop_last") %>% 
  filter(!is.na(Russia_groups_long)) %>% 
  mutate(Percentage = n / sum(n) * 100) %>% 
  kable(digits = 2)
```

Short Term Variable

```{r}
EUI_data %>% 
  filter(Year > 2022) %>% 
  group_by(Year, Russia_groups_short) %>% 
  summarise(n = n(), .groups = "drop_last") %>% 
  filter(!is.na(Russia_groups_short)) %>% 
  mutate(Percentage = n / sum(n) * 100) %>% 
  kable(digits = 2)
```

### Groups by Country

Long Term Measure

```{r}
EUI_data %>% 
  filter(Year > 2019) %>% 
  group_by(country, Russia_groups_long) %>% 
  summarise(n = n(), .groups = "drop_last") %>% 
  filter(!is.na(Russia_groups_long)) %>% 
  mutate(Percentage = n / sum(n) * 100) %>% 
  kable(digits = 2)
```

Short Term Measure

```{r}
EUI_data %>% 
  filter(Year > 2022) %>% 
  group_by(country, Russia_groups_short) %>% 
  summarise(n = n(), .groups = "drop_last") %>% 
  filter(!is.na(Russia_groups_short)) %>% 
  mutate(Percentage = n / sum(n) * 100) %>% 
  kable(digits = 2)
```


## Next steps

- Decide which measure to use for the groups.
- Decide what social medias we want to scrape to identify discourse surrounding the war in Ukraine.
  - We need API access for Meta. 
  - We can then use BERTopic to identify themes in this corpus.
  
## Dietlind's comments 

- Too many unrelated components 
- Left over group 
- What are essential vars
- 
