---
title: "Groups on Russia Progress Report"
author: "Rafael Campos-Gottardo & Alan Nemirovski"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("scripts/2_group_variables.R")
```

## Comments from meeting: Feb. 12, 2026
- Pro-Russia/Pro-Ukraine dichotomy in the long-run as opposed to doing the more granular "Russia appeaser" (only using dichotomous question around trade/security, i.e., Q73) --> appeaser group is conceptually too small over time
- "Other" group is too big
- Appeasers: also more broad appeasers (potentially on a more general note) --> sticking with smaller appeasers group for short
- Additional data sources: European Social Survey, European Values Survey (same as World Values Survey), Eurobarometer survey (look at those for addidtional Russia-related Q's?)
- Calling people that selected "Neither" or "Don't Know" as "Un-knowledgeable" (more broadly)

# Long-term (simplified) measure

Long-term measure centered on dichotomous question for trade with Russia (Pro-Russia) or defence against Russia (Pro-Ukraine); all outstanding respondents tentatively classed as "Other".

```{r}
EUI_data <- EUI_data %>% 
  mutate(ukrainegroups_long = case_when(
    #classed as Pro-Russia
    Q73 == "European countries should invest more in trade and diplomacy with Russia to improve relations" ~ "Pro-Russia",
    #classed as Pro-Ukraine
    Q73 == "European countries should invest more in defence and security to defend against Russian aggression" ~ "Pro-Ukraine",
                                       TRUE ~ "Other"))

table(EUI_data$ukrainegroups_long)

```

## Trends over time: Long-term measure

```{r}
COUNTRIES_2020 <- c("Denmark", "Finland", "France", "Germany", "Greece", "Hungary", "Italy", "Lithuania",  "Netherlands", "Poland", "Romania", "Spain", "Sweden", "UK")

#Frequencies
EUI_data %>% 
  filter(Year > 2019) %>% 
  filter(country %in% COUNTRIES_2020) %>% 
  ggplot(aes(x = ukrainegroups_long, fill = as.factor(Year), group = as.factor(Year))) +
  geom_bar(position = "dodge") +
  labs(fill = "Year", title = "Russia Attitudinal Groups: Simplified Long-Term Measure",
       x = "Groups", y = "Frequency") + 
  theme_bw()

#Proportions
EUI_data %>% 
  filter(Year > 2019) %>% 
  filter(country %in% COUNTRIES_2020) %>% 
  ggplot(aes(x = Year, fill = as.factor(ukrainegroups_long))) + 
  geom_bar(position = "fill") +
  labs(fill = "Year", title = "Russia Attitudinal Group Proportions: Simplified Long-Term Measure",
       x = "Groups", y = "Frequency") + 
  theme_bw()

```

# Short-term (simplified) measure

```{r}
#creating "short" dataset for separate short-term variable manipulation
EUI_data_short <- EUI_data %>%
  filter(Year > 2022) %>% 
  #creating counts for "proukr", "prorus", and "dk" ("don't know") camps
  mutate(Q73_proukr = as.integer(Q73 == "European countries should invest more in defence and security to defend against Russian aggression"),
         Q73_prorus = as.integer(Q73 == "European countries should invest more in trade and diplomacy with Russia to improve relations"),
         Q73_dk = as.integer(Q73 == "Neither/DK"),
         
         Q78_4_prorus = as.integer(New_Q78_4 %in% c(3, 4)),
         Q78_4_proukr = as.integer(New_Q78_4 %in% c(1, 2)),
         Q78_4_dk = as.integer(New_Q78_4 == 5),
         
         Q78_5_prorus = as.integer(New_Q78_5 %in% c(3, 4)),
         Q78_5_proukr = as.integer(New_Q78_5 %in% c(1, 2)),
         Q78_5_dk = as.integer(New_Q78_5 == 5),
         
         #summing responses in each direction per individual
         prorus_count = Q73_prorus + Q78_4_prorus + Q78_5_prorus,
         proukr_count = Q73_proukr + Q78_4_proukr + Q78_5_proukr,
         dk_count = Q73_dk + Q78_4_dk + Q78_5_dk,
         
         ukrainegroups_short = case_when(
           #at least 2/3 pro-Russia answers
           prorus_count >= 2 ~ "Pro-Russia", 
           #at least 2/3 pro-Ukraine answers
           proukr_count >=2 ~ "Pro-Ukraine",
           #exactly 1/3 pro-Russia AND 1/3 pro-Ukraine answer
           prorus_count == 1 & proukr_count == 1 ~ "Ambivalent",
           #at least 2/3 "Don't Know's"
           dk_count >= 2 ~ "Unknowledgeable",
           TRUE ~ "Other")) #"Other" included as a fail-safe category

#Every individual was classed into one of four categories
#(we can see this because there is no "Other" category in the table)
table(EUI_data_short$ukrainegroups_short)

```


## Trends over time: Short-term measure

```{r}

COUNTRIES_2023 <- c("UK", "Denmark", "Greece", "Hungary", "Lithuania", "Italy", "Poland", "Netherlands", "Romania",  "Slovakia",  "Croatia", "Bulgaria", "Spain", "Finland", "France", "Germany", "Sweden")

#Frequencies
EUI_data_short %>% 
  filter(Year > 2022) %>% 
  filter(country %in% COUNTRIES_2023) %>% 
  ggplot(aes(x = ukrainegroups_short, fill = as.factor(Year), group = as.factor(Year))) +
  geom_bar(position = "dodge") +
  labs(fill = "Year", title = "Russia Attitudinal Groups: Simplified Short-Term Measure",
       x = "Groups", y = "Frequency") + 
  theme_bw()

#Proportions
EUI_data_short %>% 
  filter(Year > 2022) %>% 
  filter(country %in% COUNTRIES_2023) %>% 
  ggplot(aes(x = Year, fill = as.factor(ukrainegroups_short))) +
  geom_bar(position = "fill") + #for frequency, change position to "dodge"
  labs(fill = "Year", title = "Russia Attitudinal Group Proportions: Simplified Short-Term Measure",
       x = "Groups", y = "Frequency") + 
  theme_bw()

```


# Country-level trends along new short-term measure

## Pro-Russia

```{r, echo = FALSE}

europe <- ne_countries(scale = "medium", returnclass = "sf") %>%
  filter(region_un == "Europe")

europe <- europe %>% 
 # st_transform(3035) %>% 
  mutate(name = replace(name, name == "United Kingdom", "UK"))

years <- c(2023, 2024, 2025)
maps_pro_Russia <- list()
for(i in 1:length(years)){
  df <- EUI_data_short %>% 
    filter(Year == years[i]) %>% 
    mutate(Pro_Russia = case_when(ukrainegroups_short == "Pro-Russia" ~ 1,
                                  TRUE ~ 0)) %>% 
    group_by(country) %>% 
    summarise(`Pro-Russia` = mean(Pro_Russia, na.rm = TRUE))
  
  
  df <- left_join(europe, df, by = c("name" = "country"))
  
  maps_pro_Russia[[as.character(years[i])]] <- df %>% 
    ggplot() + 
    geom_sf(aes(fill = `Pro-Russia`)) + 
    scale_fill_gradientn(
    colours = c("pink","darkred"), # light pink → hot pink → purple
    na.value = "grey90",
    breaks = seq(0, 0.7, 0.1),
    limits = c(0, 0.7),
    labels = scales::percent
  ) +
      coord_sf(
    xlim = c(-25, 49),   # longitude range
    ylim = c(34, 72)     # latitude range
  ) + 
    theme_void() +
      labs(fill = "% Pro-Russia",
           title = paste0(years[i])) + 
  theme(legend.position = "bottom",
        text = element_text(size = 8))
}

ggpubr::ggarrange(plotlist = maps_pro_Russia[c(1,2, 3)],
  ncol = 2,
  nrow = 2,
  common.legend = TRUE,
  legend = "bottom")

```

## Pro-Ukraine

```{r, echo = FALSE}

maps_pro_ukraine <- list()
for(i in 1:length(years)){
  df <- EUI_data_short %>% 
    filter(Year == years[i]) %>% 
    mutate(Ukraine = case_when(ukrainegroups_short == "Pro-Ukraine" ~ 1,
                                  TRUE ~ 0)) %>% 
    group_by(country) %>% 
    summarise(`Pro-Ukraine` = mean(Ukraine, na.rm = TRUE))
  
  
  df <- left_join(europe, df, by = c("name" = "country"))
  
  maps_pro_ukraine[[as.character(years[i])]] <- df %>% 
    ggplot() + 
    geom_sf(aes(fill = `Pro-Ukraine`)) + 
    scale_fill_gradientn(
    colours = c("lightblue","blue"), # light pink → hot pink → purple
    na.value = "grey90",
    breaks = seq(0, 0.7, 0.1),
    limits = c(0, 0.7),
    labels = scales::percent
  ) +
      coord_sf(
    xlim = c(-25, 49),   # longitude range
    ylim = c(34, 72)     # latitude range
  ) + 
    theme_void() +
      labs(fill = "% Pro-Ukraine",
           title = paste0(years[i])) + 
  theme(legend.position = "bottom",
        text = element_text(size = 8))
}

ggpubr::ggarrange(plotlist = maps_pro_ukraine[c(1, 2, 3)],
  ncol = 2,
  nrow = 2,
  common.legend = TRUE,
  legend = "bottom")

```

## Ambivalent

```{r, echo = FALSE}

maps_other <- list()
for(i in 1:length(years)){
  df <- EUI_data_short %>% 
    filter(Year == years[i]) %>% 
    mutate(Ambivelent = case_when(ukrainegroups_short == "Ambivalent" ~ 1,
                                  TRUE ~ 0)) %>% 
    group_by(country) %>% 
    summarise(Ambivelent = mean(Ambivelent, na.rm = TRUE))
  
  
  df <- left_join(europe, df, by = c("name" = "country"))
  
  maps_other[[as.character(years[i])]] <- df %>% 
    ggplot() + 
    geom_sf(aes(fill = Ambivelent)) + 
    scale_fill_gradientn(
    colours = c("yellow","orange"),
    na.value = "grey90",
    breaks = seq(0, 0.4, 0.1),
    limits = c(0, 0.4),
    labels = scales::percent
  ) +
      coord_sf(
    xlim = c(-25, 49),   # longitude range
    ylim = c(34, 72)     # latitude range
  ) + 
    theme_void() +
      labs(fill = "% Ambivalent",
           title = paste0(years[i])) + 
  theme(legend.position = "bottom",
        text = element_text(size = 8))
}

ggpubr::ggarrange(plotlist = maps_other[c(1, 2, 3)],
  ncol = 2,
  nrow = 2,
  common.legend = TRUE,
  legend = "bottom")

```

## Unknowledgeable

```{r, echo = FALSE}

maps_other <- list()
for(i in 1:length(years)){
  df <- EUI_data_short %>% 
    filter(Year == years[i]) %>% 
    mutate(Ambivelent = case_when(ukrainegroups_short == "Unknowledgeable" ~ 1,
                                  TRUE ~ 0)) %>% 
    group_by(country) %>% 
    summarise(Ambivelent = mean(Ambivelent, na.rm = TRUE))
  
  
  df <- left_join(europe, df, by = c("name" = "country"))
  
  maps_other[[as.character(years[i])]] <- df %>% 
    ggplot() + 
    geom_sf(aes(fill = Ambivelent)) + 
    scale_fill_gradientn(
    colours = c("lightgreen","darkgreen"),
    na.value = "grey90",
    breaks = seq(0, 0.4, 0.1),
    limits = c(0, 0.4),
    labels = scales::percent
  ) +
      coord_sf(
    xlim = c(-25, 49),   # longitude range
    ylim = c(34, 72)     # latitude range
  ) + 
    theme_void() +
      labs(fill = "% Unknowledgeable",
           title = paste0(years[i])) + 
  theme(legend.position = "bottom",
        text = element_text(size = 8))
}

ggpubr::ggarrange(plotlist = maps_other[c(1, 2, 3)],
  ncol = 2,
  nrow = 2,
  common.legend = TRUE,
  legend = "bottom")

```

