---
title: "Updated Preliminary Report"
author: "Rafael Campos-Gottardo & Alan Nemirovski"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("scripts/2_group_variables.R")
```


# New Measure of groups short term

```{r}

EUI_data_short <- EUI_data %>%
  filter(Year > 2022) 

EUI_data_short <- EUI_data_short %>% 
  bind_rows(EUI_april_2022 %>% mutate(Year = 2022))

EUI_data_short <- EUI_data_short %>% 
  mutate(Ukraine_groups = case_when(Q73 == "European countries should invest more in defence and security to defend against Russian aggression" &
                                      New_Q78_4 %in% c(3, 4) &
                                      New_Q78_5 %in% c(3, 4) ~ "Security-focused",
                                    New_Q78_4 %in% c(3, 4) &
                                      New_Q78_5 %in% c(1, 2) ~ "Conditional Ukraine Supporters",
                                    Q73 == "European countries should invest more in trade and diplomacy with Russia to improve relations" & 
                                      New_Q78_4 %in% c(1, 2) ~ "Russia collaboration",
                                    TRUE ~ "Domestic/ Distracted"),
         Ukraine_groups = factor(Ukraine_groups, levels = c("Domestic/ Distracted", "Security-focused", "Conditional Ukraine Supporters", "Russia collaboration")))
```

The Coding of the new measure

Security Focused: Q73 - Defence and Security; New_Q78_4 - Support sending military equipment; New_Q78_5 - Willing to accept higher energy costs

Conditional Supporters: New_Q78_4 - Support sending military equipment; New_Q78_5 - Not wiling to accept higher energy costs

Russian collaboration: Q73 - Trade and diplomacy; New_Q78_4 - Do not support sending military equipment

Domestic/Distracted: All those not previously categorized (The coding scheme says and don't know which is captured by all the remaining categories)

There is the code used to create the groups with spacing to see all of it
```
EUI_data_short <- EUI_data_short %>% 
  mutate(Ukraine_groups = case_when(Q73 == "European countries should invest
  more in defence and security to defend against Russian aggression" &
                                      New_Q78_4 %in% c(3, 4) &
                                      New_Q78_5 %in% c(3, 4) ~ "Security-focused",
                                    New_Q78_4 %in% c(3, 4) &
                                      New_Q78_5 %in% c(1, 2) ~ "Conditional
                                      Ukraine Supporters",
                                    Q73 == "European countries should invest 
                                    more in trade and diplomacy with Russia to
                                    improve relations" & 
                                      New_Q78_4 %in% c(1, 2) ~ 
                                      "Russia collaboration",
                                    TRUE ~ "Domestic/ Distracted"))
```
```{r}
table(EUI_data_short$Ukraine_groups)
```

### Examining the Domestic/Distracted Group

```{r}
EUI_data_short %>% 
  mutate(New_Q78_4 = case_when(New_Q78_4 %in% c(1,2) ~ "No Military Equipment",
                               New_Q78_4 %in% c(3,4) ~ "Yes Military Equipment", 
                               New_Q78_4 == 5 ~ "DK"),
         New_Q78_5 = case_when(New_Q78_5 %in% c(1,2) ~ "Do not Support Sanctions",
                               New_Q78_5 %in% c(3,4) ~ "Support Sanctions", 
                               New_Q78_5 == 5 ~ "DK"),
         Q73 = case_when(Q73 == "European countries should invest more in defence and security to defend against Russian aggression" ~ "Defence and Security",
                         Q73 == "European countries should invest more in trade and diplomacy with Russia to improve relations" ~ "Trade and Diplomacy",
                        TRUE ~ Q73)
         ) %>% 
  group_by(Q73, New_Q78_4, New_Q78_5) %>% 
  count(Ukraine_groups) %>% 
  filter(Ukraine_groups == "Domestic/ Distracted") %>% 
  ungroup() %>% 
  select(-Ukraine_groups) %>% 
  mutate(percentage = round(n/sum(n)*100, 2)) %>% 
  arrange(-percentage) %>% 
  kable(col.names = c("Plan for Russia", "Military Aid", "Sanctions", "N", "Percentage"
                      ))

```

## Trends over time: New Measure 

```{r}

COUNTRIES_2022 <- c("UK", "Denmark", "Greece", "Hungary", "Lithuania",
                    "Italy", "Poland", "Netherlands", "Romania",  "Slovakia",
                    "Croatia", "Bulgaria", "Spain", "Finland", "France",
                    "Germany", "Sweden")

#Frequencies
EUI_data_short %>% 
  filter(Year > 2021) %>% 
  filter(country %in% COUNTRIES_2022) %>% 
    mutate(Ukraine_groups = replace(Ukraine_groups, Ukraine_groups == "Conditional Ukraine Supporters", "Conditional \n Ukraine Supporters"),
      Ukraine_groups = factor(Ukraine_groups,
                                      levels = c("Security-focused",
                                                 "Conditional \n Ukraine Supporters",
                                                 "Domestic/ Distracted",
                                                 "Russia collaboration"))) %>% 
  ggplot(aes(x = Ukraine_groups, fill = as.factor(Year), group = as.factor(Year))) +
  geom_bar(position = "dodge") +
  labs(fill = "Year", title = "Russia Attitudinal Groups",
       x = "Groups", y = "Frequency") + 
  theme_bw()

EUI_data_short %>% 
  filter(Year > 2021) %>% 
  filter(country %in% COUNTRIES_2022) %>% 
  group_by(Year) %>% 
  count(Ukraine_groups) %>% 
  mutate(Percentage = n / sum(n) ) %>% 
    mutate(Ukraine_groups = replace(Ukraine_groups, Ukraine_groups == "Conditional Ukraine Supporters", "Conditional \n Ukraine Supporters"),
      Ukraine_groups = factor(Ukraine_groups,
                                      levels = c("Security-focused",
                                                 "Conditional \n Ukraine Supporters",
                                                 "Domestic/ Distracted",
                                                 "Russia collaboration"))) %>% 
    ggplot(aes(x = Ukraine_groups, y = Percentage, fill = as.factor(Year))) + 
    geom_col(position = "dodge") +
      labs(fill = "Year", title = "Russia Attitudinal Group Proportions: \n Short-Term Measure (Percentage)",
       x = "Groups", y = "Percentage") + 
    scale_y_continuous(labels = scales::percent) +
      theme_bw()
  
#Proportions
EUI_data_short %>% 
  filter(Year > 2021) %>% 
  filter(country %in% COUNTRIES_2022) %>% 
  mutate(Ukraine_groups = factor(Ukraine_groups,
                                      levels = c("Security-focused",
                                                 "Conditional Ukraine Supporters",
                                                  "Domestic/ Distracted",
                                                 "Russia collaboration"))) %>% 
  ggplot(aes(x = Year, fill = (Ukraine_groups))) +
  geom_bar(position = "fill") + #for frequency, change position to "dodge"
  labs(fill = "Year", title = "Russia Attitudinal Group Proportions: Short-Term Measure",
       x = "Groups", y = "Proportion") + 
  scale_fill_manual(values = c("darkblue", "lightblue", "forestgreen", "darkred")) + 
  theme_bw()

```

# Short Term Country Year Trends Graph 

```{r  fig.width=10, fig.height=12, out.width="\\linewidth"}
short_country_year_df <- EUI_data_short %>% 
  filter(Year > 2021) %>% 
  filter(!is.na(country)) %>% 
  mutate(Ukraine_groups = factor(Ukraine_groups,
                                      levels = c("Security-focused",
                                                 "Conditional Ukraine Supporters",
                                                 "Domestic/ Distracted",
                                                 "Russia collaboration"))) %>% 
  group_by(Year, country) %>% 
  count(Ukraine_groups) %>% 
  mutate(Percentage = n / sum(n))


countries <- short_country_year_df %>% 
  filter(Year == 2025) %>% 
  filter(Ukraine_groups == "Security-focused") %>% 
  arrange(-Percentage) %>% 
  pull(country)

short_country_year_df %>% 
  mutate(country = factor(country, levels = countries)) %>% 
    ggplot(aes(x = Year, y = Percentage, fill = as.factor(Ukraine_groups))) + 
    geom_col(position = "fill") +
      labs(fill = "Group", title = "Russia Attitudinal Group Proportions: \n Short Term Measure (by Country)",
       x = "Year", y = "Percentage",
       caption = "Ordered by Highest Percentage of Pro-Russians in 2025") + 
  facet_wrap(~country, ncol = 4) +
  scale_fill_manual(values = c("darkblue", "lightblue", "forestgreen", "darkred")) + 
    scale_y_continuous(labels = scales::percent) +
      theme_bw() + 
  theme(legend.position = "bottom") 


```

# Robustness Checks on Other measures

## Support for Aid to Ukraine

```{r}

EUI_data_short <- EUI_data_short %>% 
  mutate(Support_refugees = ifelse(New_Q78_1 %in% c(3, 4), 1, 0),
         Support_humanitarian_aid = ifelse(New_Q78_2 %in% c(3, 4), 1, 0),
         Support_EU_memebership = ifelse(New_Q78_7 %in% c(3, 4), 1, 0)
         )


EUI_data_short_long <- EUI_data_short %>% 
  pivot_longer(cols = c(Support_refugees, Support_humanitarian_aid, Support_EU_memebership),
               values_to = "Support", names_to = "Aid_type")


EUI_data_short_long %>% 
  group_by(Aid_type, Ukraine_groups) %>% 
  summarize(Mean = paste0(round(mean(Support, na.rm = TRUE) * 100, 2), "%")) %>% 
  pivot_wider(names_from = Aid_type, values_from = Mean) %>% 
  kable(col.names = c("Group", "EU Membership Ukraine", "Send Humanitarian Aid",
                      "Accept Refugees"),
        label = "Support for Ukraine Aid by Membership in Groups in Ukraine") 

```

## Preferred Outcome to the War

```{r}

EUI_data_short %>% 
    group_by(Ukraine_groups) %>% 
    count(EUI_Ukraine_Outcome) %>% 
  mutate(EUI_Ukraine_Outcome = case_match(EUI_Ukraine_Outcome,
                                          1 ~ "Russia achieves all territorial goals",
                                          2 ~ "Russia takes more territory",
                                          3 ~ "A return to the territorial situation beofre the war",
                                          4 ~ "Ukraine retakes some territory",
                                          5 ~ "Ukraine retakes all territory"),
         Ukraine_groups = factor(Ukraine_groups,
                                      levels = c("Security-focused",
                                                  "Conditional Ukraine Supporters",
                                                 "Domestic/ Distracted",
                                                 "Russia collaboration"))) %>% 
  filter(!is.na(EUI_Ukraine_Outcome)) %>% 
  mutate(n = paste0(round(n/ sum(n) * 100, 2), "%")) %>% 
  pivot_wider(names_from = Ukraine_groups, values_from = n) %>%  
  rename("Preferred Outcome" = EUI_Ukraine_Outcome) %>% 
  kable()

```

## Trust in Russia

```{r}

EUI_data_short %>% 
    group_by(Ukraine_groups) %>% 
    mutate(Trust_Russia_3 = ifelse(A5_2 >= 3, 1, 0),
           Trust_Russia_5 = ifelse(A5_2 >= 5, 1, 0),
           Trust_Russia_7 = ifelse(A5_2 >= 7, 1, 0),
           Trust_Russia_9 = ifelse(A5_2 >= 9, 1, 0)
           ) %>% 
  mutate(Ukraine_groups = factor(Ukraine_groups,
                                      levels = c("Security-focused",                      
                                                 "Conditional Ukraine Supporters",
                                                 "Domestic/ Distracted",
                                                 "Russia collaboration"))) %>% 
  filter(!is.na(EUI_Ukraine_Outcome)) %>% 
  summarize(`Percentage Trust 3` = paste0(round(mean(Trust_Russia_3, na.rm = TRUE)*100,
                                              2), "%"),
            `Percentage Trust 5` = paste0(round(mean(Trust_Russia_5, na.rm = TRUE)*100,
                                              2), "%"),
            `Percentage Trust 7` = paste0(round(mean(Trust_Russia_7, na.rm = TRUE)*100,
                                              2), "%"),
            `Percentage Trust 9` = paste0(round(mean(Trust_Russia_9, na.rm = TRUE)*100,
                                              2), "%"),
            ) %>% 
  rename(Group = Ukraine_groups) %>% 
  kable()

```

## Trust in Ukraine

```{r}
EUI_data_short %>% 
    group_by(Ukraine_groups) %>% 
    mutate(Trust_Russia_3 = ifelse(A5_3 >= 3, 1, 0),
           Trust_Russia_5 = ifelse(A5_3 >= 5, 1, 0),
           Trust_Russia_7 = ifelse(A5_3 >= 7, 1, 0),
           Trust_Russia_9 = ifelse(A5_3 >= 9, 1, 0)
           ) %>% 
  mutate(Ukraine_groups = factor(Ukraine_groups,
                                      levels = c("Security-focused",
                                                 "Conditional Ukraine Supporters",
                                                 "Domestic/ Distracted",
                                                 "Russia collaboration"))) %>% 
  filter(!is.na(EUI_Ukraine_Outcome)) %>% 
  summarize(`Percentage Trust 3` = paste0(round(mean(Trust_Russia_3, na.rm = TRUE)*100,
                                              2), "%"),
            `Percentage Trust 5` = paste0(round(mean(Trust_Russia_5, na.rm = TRUE)*100,
                                              2), "%"),
            `Percentage Trust 7` = paste0(round(mean(Trust_Russia_7, na.rm = TRUE)*100,
                                              2), "%"),
            `Percentage Trust 9` = paste0(round(mean(Trust_Russia_9, na.rm = TRUE)*100,
                                              2), "%"),
            ) %>% 
  rename(Group = Ukraine_groups) %>% 
  kable()
```


## Greatest Threat

```{r}
EUI_data_short %>% 
    group_by(Ukraine_groups) %>% 
    count(Q68) %>% 
  mutate(Q68 = case_match(Q68,
                          1 ~ "Terrorism",
                          2 ~ "Russia",
                          3 ~ "The US",
                          4 ~ "China",
                          5 ~ "Nuclear Proliferation",
                          6 ~ "Other",
                          7 ~ "Don't Know"
                          ),
         Ukraine_groups = factor(Ukraine_groups,
                                      levels = c("Security-focused",
                                                 "Conditional Ukraine Supporters",
                                                "Domestic/ Distracted",
                                                 "Russia collaboration"))) %>% 
  mutate(n = paste0(round(n/ sum(n) * 100, 2), "%")) %>% 
  pivot_wider(names_from = Ukraine_groups, values_from = n) %>%  
  rename(`Biggest Threat` = Q68) %>% 
  kable()

```

## Responsibility for the Conflict

```{r}

EUI_data_short %>% 
    group_by(Ukraine_groups) %>% 
    count(New_Q79) %>% 
  mutate(New_Q79 = case_match(New_Q79,
                          1 ~ "Entirely NATO",
                          2 ~ "More NATO Than Russia",
                          3 ~ "NATO and Russia Equally",
                          4 ~ "More Russia than NATO",
                          5 ~ "Entirely Russia",
                          6 ~ "Don't Know"
                          ),
         Ukraine_groups = factor(Ukraine_groups,
                                      levels = c("Security-focused",
                                                 "Conditional Ukraine Supporters",
                                                "Domestic/ Distracted",
                                                 "Russia collaboration"))) %>% 
  filter(!is.na(New_Q79)) %>% 
  mutate(n = paste0(round(n/ sum(n) * 100, 2), "%")) %>% 
  pivot_wider(names_from = Ukraine_groups, values_from = n) %>%
  rename("Responsibilty for War" = New_Q79) %>% 
  kable()

```

## Threat of Nuclear Escalation Should be considered when making aid decisions

```{r}
EUI_data_short %>% 
    group_by(Ukraine_groups) %>% 
  mutate(
         Ukraine_groups = factor(Ukraine_groups,
                                      levels = c("Security-focused", 
                                                 "Conditional Ukraine Supporters",
                                                "Domestic/ Distracted",
                                                 "Russia collaboration"))) %>% 
summarize(`Threat of Escalation` = mean(ForeignPolicyDecisions2_2, na.rm = TRUE)) %>% 
  mutate(`Threat of Escalation` = paste0(round(`Threat of Escalation` * 100, 2), "%")) %>% 
  kable()
```

